Structural replication of nanoporous media using procedural noise

### 2017-02-05 ms.tex
Smallies

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate geometries at arbitrary scales. In addition, this method is extremely fast and low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

\begin{figure}
\includegraphics[width=.45\textwidth]{gofr_plot.png}
\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). \todo{Bytt denne figuren med VMD sin beregning. Normaliseringen vår er gal.}}
\label{fig:gofr1}
\end{figure}
Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
\begin{figure}
\includegraphics[width=.45\textwidth]{comparison.png}
\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO$_2$ input data set used for parameter estimation. }
\label{fig:porous_vs_model}
\end{figure}

The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area of the original data set was $\sim 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging between $\sim 28 000 \AA^2$ and $\sim 30 000 \AA^2$.
\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate geometries at arbitrary scales. In addition, this method is extremely fast and has low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

\begin{figure}
\includegraphics[width=.45\textwidth]{gofr_plot.png}
\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). \todo{Bytt denne figuren med VMD sin beregning. Normaliseringen vår er gal.}}
\label{fig:gofr1}
\end{figure}
Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
\begin{figure}
\includegraphics[width=.45\textwidth]{comparison.png}
\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO$_2$ input data set used for parameter estimation. Although the pore geometry seems to be similar, we see some artifacts, such as single loose atoms in the generated geometry.}
\label{fig:porous_vs_model}
\end{figure}

The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}
